---
title: "Creative Assignment 1"
author: "Maggie Weese"
date: "9/9/2020"
output: 
  html_document:
      toc: true
      toc_float: true
---
# Loading Library and Datasets

Below is my code for loading my libraries and variables. I looked at data from the 2010 census at the tract level. The variables I selected include:

1. Percentage of rental households
2. Majority race in census tract
3. Majority sex
4. Median age
5. Average household size
6. Percentage of "Nonfamily" households


```{r load libraries}
library(ggplot2)
library(tidyverse)
library(tidycensus)
library(RColorBrewer)
library(wesanderson)

```
```{r loading data varibles}
#The code for variables rent, majority race, and age are used from Carole's demo data on GitHub

vara <- c(tot_hh = "H001001",
         tot_rent_hh = "H004004",
         avg_hh_size = "H012001", 
         tot_pop = "P003001",
         nh_white_pop = "P005003",
         nh_black_pop = "P005004",
         nh_asian_pop = "P005006",
         hispanic_pop = "P005010",
         total_hh = "PCT016001",
         total_nonfam_hh = "PCT016021",
         total_sex = "P012001",
         total_fem = "P012002",
         total_male = "P012026",
         med_age = "P013001")

Bmore_tracts <- get_decennial(geography = "tract", variables = vara,
                           state = "MD", county = "Baltimore City",
                           output = "wide") %>%
  mutate(pct_rental = tot_rent_hh / tot_hh,
         pct_nonfam_hh = total_nonfam_hh / total_hh,
         majority_sex = case_when(
           total_fem / total_sex > 0.5 ~ "fem",
           total_male / total_sex > 0.5 ~ "male"),
         majority_race_eth = case_when(
           nh_white_pop / tot_pop > 0.5 ~ "nh_white",
           nh_black_pop / tot_pop > 0.5 ~ "nh_black",
           nh_asian_pop / tot_pop > 0.5 ~ "nh_asian",
           hispanic_pop / tot_pop > 0.5 ~ "hispanic",
           (nh_white_pop + nh_black_pop + nh_asian_pop + hispanic_pop) /
             tot_pop < 0.5 ~ "other",
           TRUE ~ "None")) %>%
  filter(tot_pop > 0, tot_hh > 0) %>%
  select(GEOID, majority_race_eth, 
         pct_rental, avg_hh_size, tot_pop, med_age, majority_sex, pct_nonfam_hh) 

```
 
## Creation of Graphs

### Graph 1 - Scatter Plot - Iteration 1

In this initial scatter plot I looked at average household size in each Baltimore census tracts compared to the percentage of rental households in the tract. I added color to designate the majority gender in each tract. While a good initial try, this graph did not give me a very clear picture of any relationship and I think the use of color to highlight sex may take away from/abstract the relationship. 

The basis of this code came from Carole's example in Github.

```{r scatter plot1}
ggplot(Bmore_tracts, 
       aes(x = pct_rental,
           y = avg_hh_size,
           color = majority_sex)) +
  geom_point() +
  scale_x_continuous(name = "Percent of rental households",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Average household size",
                     breaks = seq(10, 80, by = 10)) +
  scale_color_discrete(name = "Majority Sex", 
                       labels = c("Female", "Male")) +
  theme_bw()
```

### Graph 2 - Scatter Plot - Iteration 2

As we discussed in class, often less is more in data visualization so I removed the color-coding and instead added a trend line. While still not great, I think this graph does a better job highlighting the relationship between the variables. 

The basis of this code came from Carole's example in Github.

```{r scatter plot - version 2}
ggplot(Bmore_tracts, 
       aes(x = pct_rental,
           y = avg_hh_size)) +
  geom_point() +
  stat_smooth(method = "lm") +
  scale_x_continuous(name = "Percent of rental households",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Average household size", breaks = seq(0, 5, by = 1)) +
  theme_bw()
```

### Graph 3 - Scatter Plot, Use of Size 

In this plot, I played around with the use of dot size to represent a variable and learned from R's error code that it is not advised to use size for a categorical set of data.

```{r scatter plot3}

ggplot(Bmore_tracts, 
       aes(x = pct_rental,
           y = med_age,
           color = majority_sex,
           size = majority_race_eth)) +
  geom_point() +
  scale_x_continuous(name = "Percent of rental households",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Median age",
                     breaks = seq(10, 80, by = 10)) +
  scale_color_discrete(name = "Majority Sex", 
                       labels = c("Female", "Male")) +
  scale_size_discrete(name = "Majority Race/Ethnicity",
          labels = c("Non-Hispanic-Black",
                                  "Non-Hispanic White",
                                  "No Majority")) +
  theme_bw()
```

### Graph 4 - Violin Plots and Exploring Fill

In this plot I used violin plots to understand the relationship between the percentage of rental households and majority race. While fun to make, I did not think this graph told too much except that the percentage of rental households is relatively evenly distributed actross the three race/ethnicities.

The basis of this code came from Carole's example in Github with modifications.

```{r violin plot}
ggplot(Bmore_tracts, 
       aes(x = majority_race_eth,
           y = pct_rental,
           fill= majority_race_eth)) +
  geom_violin() +
  scale_fill_manual(values = wes_palette("GrandBudapest2", n = 3)) +
  scale_x_discrete(name = "Majority race/ethnicity", 
                            labels = c("Non-Hispanic-White",
                                  "Non-Hispanic White",
                                  "No Majority")) +
  scale_y_continuous(name = "Percentage of rental households",
                    breaks = seq(0,1 , by = 0.1), 
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  theme_bw() +
  theme(legend.position = "none")
```

### Graph 5 - Violin Plots

I then went on to see the different ways I could represent the relationship between categorical and continuous data, focusing on the relationship between majority sex and percentage of rental households. 

The basis of this code came from Carole's example in Github with modifications. 
```{r violin plot2}
ggplot(Bmore_tracts, 
       aes(x = majority_sex,
           y = pct_rental,
           color = majority_sex)) +
  geom_violin() +
  scale_fill_manual(values = "Blue") +
  scale_x_discrete(name = "Majority Sex",
                   labels = c("Male",
                                  "Female")) +
  scale_y_continuous(name = "Percentage of rental households",
                     breaks = seq(0,1 , by = 0.1), 
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  theme_bw() +
  theme(legend.position = "none")
```

### Graph 6 - Boxplots

My second interaction of representing the relationship between majority sex and percentage of rental households was through boxplots. I found these graphs to be much clearer to read and understand.  

The basis of this code came from Carole's example in Github with modifications. 

```{r boxplot}
ggplot(Bmore_tracts, 
       aes(x = majority_sex,
           y = pct_rental,
           color = majority_sex)) +
  geom_boxplot() +
  scale_x_discrete(name = "Majority sex",
                       labels = c("Male",
                                  "Female")) +
  scale_y_continuous(name = "Percentage of rental households",  
                     breaks = seq(0,1 , by = 0.1), 
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  theme_bw() +
  theme(legend.position = "none") 

```

### Graph 7 - Categorical Data through Bar Graphs - Iteration 1

In this graph I explored representing two pieces of categorical data - majority race/ethnicity and majority sex. I made two iterations of this graph, one that stacked the data and one that showed the data side by side. While I thought I would be partial to showing them side-by-side, I actually like the stacked representation better and think it is more clear.  

The basis of this code came from Carole's example in Github.

```{r bar graph1}
ggplot(Bmore_tracts, 
       aes(x = majority_race_eth, fill = majority_sex)) +
  geom_bar(position = "dodge") +
  scale_y_continuous(name = "Number of tracts") +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Non-Hispanic Black", "Non-Hispanic White", "No Majority")) +
    scale_fill_discrete(name = "Majority Sex", 
                       labels = c("Female", "Male"))+
  theme_bw() 
```

### Graph 8 - Categorical Data through Bar Graphs - Iteration 2

The basis of this code came from Carole's example in Github.

```{r bar graph2}
ggplot(Bmore_tracts, 
       aes(x = majority_race_eth, fill = majority_sex)) +
  geom_bar() +
  scale_y_continuous(name = "Number of tracts") +
  scale_x_discrete(name = "Majority race/ethnicity",
                       labels = c("Non-Hispanic Black", "Non-Hispanic White", "No Majority")) +
    scale_fill_discrete(name = "Majority Sex", 
                       labels = c("Female", "Male"))+
  theme_bw() 
```

### Graph 9 Scatter Plots with three Variables - Iteration 1

In this graph I explored the relationship between median age, percentage of households considered non-family and majority race/ethnicity. Similar to above, I did not find this graph to be the most helpful with all three variables so did a second iteration of it with a line of best fit for majority race/ethnicity. 

The basis of this code came from Carole's example in Github.

```{r scatter plot4}
ggplot(Bmore_tracts, 
       aes(x = pct_nonfam_hh,
           y = med_age,
           color = majority_race_eth)) +
  geom_point() +
  scale_x_continuous(name = "Percentage of households condsidered non-family",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Median age within tract",
                     breaks = seq(10, 80, by = 10)) +
  scale_color_discrete(name = "Majority race/ethnicity",
                       labels = c("Non-Hispanic Black",
                                  "Non-Hispanic White",
                                  "No majority")) +
  theme_bw() 
```

### Graph 10 Scatter Plots with three Variables - Iteration 2

The line of best fit in this graph helps to highlight the relationship between median age and percentage of households considered non-family for each race/ethnicity - a relationship that was hard to see in the last graph.  that was otherwise kind of abstract and hard to see. 

The basis of this code came from Carole's example in Github.

```{r scatter plot with fitted line}
ggplot(Bmore_tracts, 
       aes(x = pct_nonfam_hh,
           y = med_age,
           color = majority_race_eth)) +
  geom_point() +
  stat_smooth(method = "lm") +
  scale_x_continuous(name = "Percentage of households condsidered non-family",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Median age within tract",
                     breaks = seq(10, 80, by = 10)) +
  scale_color_discrete(name = "Majority race/ethnicity",
                       labels = c("Non-Hispanic Black",
                                  "Non-Hispanic White",
                                  "No majority")) +
  theme_bw() 
```

### Graph 11 - Bar Graph of Continuous Data

This graph shows the distribution in median age across all tracts. I am not sure why but I find this graph really interesting although not the most visually appealing of the bunch. 

```{r histogram graph}
ggplot(Bmore_tracts, aes(x = med_age, n = 3)) +
  scale_fill_manual(values = c("blue")) +
  scale_x_continuous(name = "Median age within tract",
                     expand = c(0,0)) +
  scale_y_continuous(name = "Number of tracts", expand = c(0,0), breaks = seq(0, 80, by = 10)) +
  geom_histogram(binwidth = 5, color = "black", fill = "orange") + 
  theme_bw()
```

