---
title: "Creative Assignment 2"
author: "Maggie Weese"
date: "9/18/2020"
output: 
  html_document:
    theme: spacelab
    toc: true
    toc_float: true
    code_folding: hide
 
---
# Creating Maps, Projection Coordinate Systems and Mapping Variables in R Studio

## Accessing Libraries

```{r, message=FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(tidyverse)
library(ggspatial)
library(cartogram)
library(ggthemes)
library(ggrepel)
library(ggrepel)
```

## Projections
### World Map

In this section I explore pre-set projections. First, I created a variable named world to create a world map. The code was based on the code provided in the online tutorial by Carole (thank you!).
```{r}
world <- ne_countries(scale = 50, returnclass = "sf") 

world_map <- ggplot(world) +
  geom_sf() +
  theme_bw()

world_map
```

### South America
Next, I decided to look more closely at South America and created a variable named SA_countries to map South America. The code was based on the code provided in the online tutorial by Carole (thank you!).

```{r}
SA_countries <- world %>%
  filter(continent == "South America")

SA_map <- ggplot(SA_countries) +
  geom_sf() + 
  theme_bw()

SA_map
```

## Adding Alternative Coordinate Reference Systems

In this section I explore changing coordinate reference systems for the map of South American. I added three alternative coordinate reference systems from the site Spatial Reference, including the Brazil Polyconic Zeroed False Northing and Easting (SA_Brazil), ICN Regional (SA_ICN) and the Bicontinental stereographic projection for the Argentine Republic (SA_Argen).

```{r}

SA_Brazil <- "+proj=poly +lat_0=0 +lon_0=-54 +x_0=0 +y_0=0 +ellps=aust_SA +units=m +no_defs"
SA_ICN <- "+proj=lcc +lat_1=9 +lat_2=3 +lat_0=6 +lon_0=-66 +x_0=1000000 +y_0=1000000 +ellps=intl +units=m +no_defs "
SA_MX <- "+proj=aea +lat_1=14.5 +lat_2=32.5 +lat_0=24 +lon_0=-105 +x_0=0 +y_0=0 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"

```


### Projection 1 - Brazil Polyconic Zeroed False Northing and Easting
For this map, the coordinate reference system was changed to the Brazil Polyconic Zeroed False Northing and Easting projection.
```{r} 

ggplot(SA_countries) +
  geom_sf() + 
  coord_sf(crs = SA_Brazil) +
  ggtitle("Map of South America - Brazil Polyconic Projection") + 
  annotation_scale() + 
  theme_bw()

```

### Projection 2 - Mexico Albers Equal Area Conic
For this map, the coordinate reference system was changed to the Mexico Albers Equal Area Conic projection.
```{r}

ggplot(SA_countries) +
  geom_sf() + 
  coord_sf(crs = SA_MX) +
  ggtitle("Map of South America - Mexico Albers Equal Area Conic") + 
  theme_bw()

```

### Projection 3 - ICN Regional
For this map, the coordinate reference system was changed to the ICN Regional projection.
```{r, message=FALSE}

ggplot(SA_countries) +
  geom_sf() + 
  coord_sf(crs = SA_ICN) +
  ggtitle("Map of South America - ICN Regional Projection") + 
  annotation_scale() + 
  theme_bw() 


```

## Mapping Variables 
I decided to use the map of South America with the Brazil Polyconic Projection in the next section of the assignment to map a variable spatially.

```{r}
sa_transformed <- SA_countries %>%
  st_transform(crs = SA_Brazil)

ggplot(sa_transformed) +
  geom_sf() +
  ggtitle("Map of South America - Brazil Polyconic Projection") + 
  annotation_scale() + 
  theme_bw()
```

### Downloading COVID-19 Data
I downloaded country level COVID-19 data from John Hopkins University’s Center for Systems Science and Engineering - https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_daily_reports

```{r, message=FALSE}
covid_SA<- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-19-2020.csv")

sum_covid <- covid_SA %>% group_by(Country_Region) %>% summarize(newDeath = sum(Deaths))

```

### Chloropleth Map of COVID 19 Deaths in South America as of September 19 2020
Here I create a Chloropleth map. The code is based on code from the online tutorial by Carole. Alex also provided support on code to create labels and combine data on death for each country (thank you!). One challenge I had was preventing the overlapping of the country labels. I found a package called ggrepel that prevents label overlap within geom_text but does not seem to work with geom_sf_text.

```{r, warning=FALSE, fig.width=8, fig.height=8}

covid_countries <-sa_transformed %>%
  left_join(sum_covid, by = c("name" = "Country_Region"))

covid_countries3 <- covid_countries[-c(7),]


ggplot(covid_countries3, aes(fill = newDeath, label = "Country_Region")) +
  geom_sf(color = "gray") +
  scale_fill_continuous_tableau(palette = "Orange",
    name = "Number of\nConfirmed\nCOVID-19\nDeaths",
    breaks = seq(0, 150000, by = 25000),
    labels = formatC(seq(0, 150000, by = 25000), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
   ggtitle("Confirmed COVID-19 Deaths in South America as of 19 September 2020") +
  theme(legend.position = "right",
        legend.direction="vertical") +
  geom_sf_text(aes(label = name), size=4, color = "gray10", inherit.aes = FALSE) +
  annotation_scale() + 
  annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
  theme(plot.title = element_text(face="bold", size=12, hjust = 0.5))


```

```{r, warning=FALSE, fig.width=8, fig.height=8}
covid_countries3 <-sa_transformed %>%
  left_join(sum_covid, by = c("name" = "Country_Region"))

covid_countries3 <- covid_countries[-c(7),]

ggplot(covid_countries3, aes(fill = newDeath, label = "Country_Region")) +
  geom_sf(color = NA) +
  scale_fill_viridis_c(option = "viridis", direction = -1,
    name = "Number of\nConfirmed\nCOVID-19\nDeaths",
    breaks = seq(0, 150000, by = 25000),
    labels = formatC(seq(0, 150000, by = 25000), 
                     big.mark = ",", format = "f", digits = 0)) +
  theme_map() +
  ggtitle("Confirmed COVID-19 Deaths in South America as of 19 September 2020") +
  theme(legend.position = "right",
        legend.direction="vertical") +
   geom_sf_text(aes(label = name), size=3.5, color = "slateblue", fontface = "bold", inherit.aes = FALSE) +
  annotation_scale() + 
  annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
  theme(plot.title = element_text(face="bold", size=12, hjust = 0.5))

 
```

### Proportional Symbol Map of COVID 19 Deaths in South America as of September 19 2020
Here I create a proportional symbol map. The code is based on code from the online tutorial by Carole (thank you!). 

```{r, warning=FALSE, fig.width=8, fig.height=8}
covid_circles <- covid_countries %>%
  st_centroid()

ggplot(covid_countries) +
  geom_sf(fill = NA, color = "gray") +
  geom_sf(data = covid_circles, 
          aes(size = newDeath),
          alpha = 0.7, color = "darkorange3") +
    scale_size_continuous(name = "Number of Confirmed\nCOVID-19 Deaths",
    breaks = seq(0, 30000, by = 5000),
    labels = formatC(seq(0, 30000, by = 5000),
                     big.mark = ",", format = "f", digits = 0),
    range = c(0, 20)) +
  ggtitle("Confirmed COVID-19 Deaths in South America as of 19 September 2020") +
  annotation_scale() + 
  annotation_north_arrow(location = "bl", which_north = "true", 
        pad_x = unit(0.25, "in"), pad_y = unit(0.5, "in"),
        style = north_arrow_fancy_orienteering) +
  theme(plot.title = element_text(face="bold", size=12, hjust = 0.5)) +
  theme_void() 


```

### Dorling Cartogram of COVID 19 Deaths in South America as of September 19 2020
Here I created a dorling cartogram. The code is based on the code from the online tutorial by Carole. Alex also provided support on code for addressing all countries with deaths labeled N/A (in this case, the Falkland Islands). Thank you both!

```{r, warning=FALSE, fig.width=8, fig.height=8}

covid_countries3[is.na(covid_countries3)] <- 0

covid_countries3 <- covid_countries[-c(7),]

covid_dorling <- covid_countries3 %>%
 cartogram_dorling("newDeath")

ggplot(covid_dorling, aes(fill = newDeath)) +
 geom_sf(color = NA) +
 scale_fill_gradient2_tableau(
    name = "Number of\nConfirmed\nCOVID-19\nDeaths",
    breaks = seq(0, 150000, by = 25000),
    labels = formatC(seq(0, 150000, by = 25000), 
                     big.mark = ",", format = "f", digits = 0)) +
  ggtitle("Confirmed COVID-19 Deaths in South America as of 19 September 2020") +
  theme(plot.title = element_text(face="bold", hjust = 0.5))+
  geom_sf_label(aes(geometry, label = name), size = 3.5, color = "gray20", fill = NA, fontface = "bold", label.size =   NA) +
  theme_map() +
  theme(legend.background = element_blank()) +
  theme(plot.title = element_text(face="bold", size=12, hjust = 0.5)) +
  theme(legend.position = "right",
        legend.direction="vertical") 

```

### Modified Area Map of COVID 19 Deaths in South America as of September 19 2020
Here I create a proportional symbol map. The code is based on code from the online tutorial by Carole (thank you!). 
```{r, warning=FALSE, message=FALSE, error=FALSE, fig.width=8, fig.height=8}

covid_countries3 <- covid_countries[-c(7),]

covid_cart_cont <- covid_countries3 %>% 
  cartogram_cont("newDeath")

ggplot(covid_cart_cont, aes(fill = newDeath)) +
 geom_sf(color = "gray") +
 scale_fill_gradient2_tableau(
    name = "Number of\nConfirmed\nCOVID-19\nDeaths",
    breaks = seq(0, 150000, by = 25000),
    labels = formatC(seq(0, 150000, by = 25000), 
                     big.mark = ",", format = "f", digits = 0)) +
  ggtitle("Confirmed COVID-19 Deaths in South America as of 19 Sept 2020") + 
  geom_sf_label(aes(geometry, label = name), size = 3.5, color = "gray20", fill = NA, label.size =   NA, fontface = "bold") +
  theme_map() +
  theme(legend.background = element_blank())+
  theme(plot.title = element_text(face="bold", hjust = 0.5)) +
  theme(legend.position = "right",
        legend.direction="vertical") 

```

```{r, warning=FALSE, message=FALSE, error=FALSE, fig.width=8, fig.height=8}

covid_countries3 <- covid_countries[-c(7),]

covid_cart_ncont <- covid_countries3 %>% 
  cartogram_ncont("newDeath")

ggplot(covid_cart_ncont, aes(fill = newDeath)) +
 geom_sf(color = "gray") +
 scale_fill_gradient2_tableau(
    name = "Number of\nConfirmed\nCOVID-19\nDeaths",
    breaks = seq(0, 150000, by = 25000),
    labels = formatC(seq(0, 150000, by = 25000), 
                     big.mark = ",", format = "f", digits = 0)) +
  ggtitle("Confirmed COVID-19 Deaths in South America as of 19 Sept 2020") +
  geom_sf_label(aes(geometry, label = name), size = 3.5, color = "gray20", fill = NA, label.size =   NA, fontface = "bold") +
  theme_map() +
  theme(axis.title.x = element_text(size = 0), axis.title.y = element_text(size = 0)) + 
  theme(plot.title = element_text(face="bold", hjust = 0.5)) +
  theme(legend.position = "right",
        legend.direction="vertical") 


```


